---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load_process_data, echo = TRUE}
# Downloading and uploading data
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)

activity <- read.csv(unzip(temp,"activity.csv"))
unlink(temp)

activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r mean_steps, echo = TRUE}

# Load package for group_by function
library(dplyr)

# Calculate the total number of steps taken per day, excludind NAs
steps <- activity[!is.na(activity),] %>% group_by(date) %>% summarise(total_per_day = sum(steps))

# Clear NAs
steps <- steps %>% na.omit()

# use base plotting system and more bins than the default setting
hist(steps$total_per_day, breaks = 20, 
     main = paste("Total Number of Steps Taken Each Day\n",
                  "Mean Steps per day =",round(mean(steps$total_per_day),2),"\n",
                  "Median Steps =", median(steps$total_per_day)),
     col = "black", border = "white", xlab = "Step", axes = FALSE)
axis(1)
axis(2, las = 1)

```


## What is the average daily activity pattern?
Time series plot of the average number of steps taken

``` {r daily_activity,  echo = TRUE}
library(stats)
avg_step <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(avg_step$interval, avg_step$steps, type = "l", lwd = 2, col = "navy",
     main = paste("Time Series: Average Number of Steps Taken \nMaximum average of steps happens at",avg_step$interval[which.max(avg_step$steps)],"th interval"),
                  , axes = FALSE,
     xlab = "5-minute interval", ylab = "Average number of steps")
axis(1)
axis(2, las = 1)

```

## Imputing missing values
Calculate the number of missing values

``` {r missing_values, echo = TRUE}
sum(is.na(activity))
```

As there are 2304 NAs, this code will implement 5-min average to fill these gaps in database.

``` {r NAs_replacement, echo = TRUE}
df <- activity # new dataset called imp
for (i in avg_step$interval) {
    df[df$interval == i & is.na(df$steps), ]$steps <- avg_step$steps[avg_step$interval == i]
}
```

``` {r hist_new_df, echo = TRUE}
steps <- df %>% group_by(date) %>% summarise(total_per_day = sum(steps))

# Clear NAs
steps <- steps %>% na.omit()

# use base plotting system and more bins than the default setting
hist(steps$total_per_day, breaks = 20, 
     main = paste("Total Number of Steps Taken Each Day\n",
                  "Mean Steps per day =",round(mean(steps$total_per_day),2),"\n",
                  "Median Steps =", round(median(steps$total_per_day),2)),
     col = "black", border = "white", xlab = "Step", axes = FALSE)
axis(1)
axis(2, las = 1)

```

## Are there differences in activity patterns between weekdays and weekends?

``` {r weekend_vs_weedays,  echo = TRUE}
df$day <- weekdays(df$date)
df$week <- ""
df[df$day == "sábado" | df$day == "domingo", ]$week <- "weekend"
df[!(df$day == "sábado" | df$day == "domingo"), ]$week <- "weekday"
df$week <- factor(df$week)

avg_step_df <- aggregate(steps ~ interval + week, data = df, mean)

library(lattice)
xyplot(steps ~ interval | week, data = avg_step_df, type = "l", lwd = 2,
       layout = c(1, 2), 
       xlab = "5-minute interval", 
       ylab = "Average number of steps",
       main = "Average Number of Steps Taken (across all weekday days or weekend days)")
```